---
title: "lab 5"
author: "Tor Fibiger"
date: "10/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Lab 5 objectives
- Learn to parse data with 'lubridate'
- Get counts of observatoins with count()
- Uncount() rows
- One and two sample t tests
- Create a heatmap with geom_tile()

```{r, inlude = FALSE}
# Read in data and attach packages

library(tidyverse)
library(here)
library(janitor)

lobster_abundance <- read_csv(here("data", "lobster_abundance.csv"),
                              na = "-99999") %>% 
  clean_names()

```

Use the tidyr::uncount() function to convert our lobster data from frequency format to case format

```{r}
lobster_tidy <- lobster_abundance %>% 
  uncount(lobster_count)
```

#### exploratory data visualization

were only going to consider 'site' as our variable of interest

```{r}
ggplot(data = lobster_tidy, aes(x = site, y = size_mm)) +
  geom_jitter(aes(color = site), alpha = 0.5, width = 0.2)

#Histogram

ggplot(data = lobster_tidy, aes(x = size_mm)) +
  geom_histogram(aes(fill = site)) +
  facet_wrap(~site, scales = "free")

ggplot(lobster_tidy, aes(sample = size_mm)) +
  geom_qq() +
  facet_wrap(~site)


```

#### Convert date column to class 'Date'

Use lubridate package to convert to date formate, and then to help us easily parse month and year

```{r}
lobster_date <- lobster_tidy %>% 
  mutate(date_new = lubridate::mdy(date))
```

Now let's parse year and month using lubridate::month() and lubridate::year()

```{r}
lobster_parse_date <- lobster_date %>% 
  mutate(obs_month = lubridate::month(date_new, label = TRUE),
         obs_year = lubridate::year(date_new))
```

Now lets find counts of observed lobsters based on different hierarchical groupings

First: lets count lobsters by year and month

```{r}
lobster_ym <- lobster_parse_date %>% 
  dplyr::count(obs_year, obs_month)

lobster_y <- lobster_parse_date %>% 
  dplyr::count(obs_year)

lobster_site <- lobster_parse_date %>% 
  dplyr::count(site)
```

If we want to create a summary table that contains stats other than counts by group, its easier to use group_by() + n()

```{r}
lobster_summary <- lobster_parse_date %>% 
  group_by(site) %>% 
  summarise(mean_size = mean(size_mm, na.rm = TRUE),
            sd_size = sd(size_mm, na.rm = TRUE),
            lobster_number = n())
```

#### find confidence intervals

Use t.test() function to find confidence intervals (for one sample) and perfrom t testst ons hte htwotjeo different samles

```{r}
ivee_lobster <- lobster_tidy %>% 
  filter(site == "IVEE") %>% 
  pull(size_mm)

t.test(ivee_lobster)
```

#### two sample t test to compare means

we want to ask is there a significant difference in lobster lengths at Naples and Mohawk reefs?

weve done our necessary exploratory analyses to determine that a 2-sample t-test for means comparison is appropriate.

```{r}
naple_sample <- lobster_tidy %>% 
  filter(site == "NAPL") %>% 
  pull(size_mm)

mohk_sample <- lobster_tidy %>% 
  filter(site == "MOHK") %>% 
  pull(size_mm)

mn_ttest <- t.test(mohk_sample, naple_sample)
```

There is a significant difference in lobster lengths between Naples and Mohawk Reef

```{r}
lobster_mn <- lobster_tidy %>% 
  filter(site %in% c("NAPL", "MOHK"))

mn_ttest2 <- t.test(size_mm ~ site, lobster_mn)


mn_ttest2$p.value
mn_ttest2$statistic

```

Here is the p-value: `r mn_ttest2$p.value`

here is my t-statistic: `r mn_ttest2$statistic`


#### Now a heat map

```{r}
lobster_ys <- lobster_parse_date %>% 
  count(obs_year, site)

ggplot(data = lobster_ys, aes(x = obs_year, y = site)) +
  geom_tile(aes(fill = n))
```

